## Load Data

```{r}
datasetRAW = read.csv(file="dpc-covid19-ita-regioni.csv")
colorsRaw = read.csv(file="covid-19-zone.csv")
```

## Prepare Data



#### Select the relevant columns only

```{r}


dataset = subset(datasetRAW, denominazione_regione=="Lombardia", select=c(data, ricoverati_con_sintomi, terapia_intensiva, totale_ospedalizzati, nuovi_positivi, deceduti, tamponi))

colors = subset(colorsRaw, select=c(Giorno,Lombardia))

```

#### Rename columns

```{r}
names(colors)=c("date","colorLabel")
names(dataset) = c("date","hospitalized_with_symptoms","intensive_care","total_hosp","new_positives","deaths","tests")


```


#### Deaths and tests per day

```{r}
deaths_per_day = rep(0,length(dataset$deaths))
tests_per_day = rep(0,length(dataset$deaths))
for(i in 2:length(dataset$deaths)){
  deaths_per_day[i] = dataset$deaths[i] - dataset$deaths[i-1]
  tests_per_day[i] = dataset$tests[i] - dataset$tests[i-1]
}
dataset = cbind(dataset,deaths_per_day,tests_per_day )


```

#### Date type for date

```{r}
dataset$date = as.Date( substr(dataset$date,0,10),format="%Y-%m-%d")
colors$date = as.Date( colors$date, format="%Y-%m-%d")
```

#### Add day of the week
```{r}
dataset$week_day = weekdays(dataset$date)
```

#### Time interval

From 2020-12-06 to 2021-07-05

```{r}
start = as.Date("2020-12-06")
end = as.Date("2021-07-05")
```

#### Get the average number of detected positives over the previous 7 days

```{r}
newpos_av7D = c()
hosp_av7D = c()
intensive_care_av7D = c()
hospitalized_with_symptoms_av7D = c()
day = start
while(day <= end){
  newpos_av7D = append(newpos_av7D,mean(dataset$new_positives[dataset$date>=day-7 & dataset$date <= day-1]))
  
  hosp_av7D = append(hosp_av7D,mean(dataset$total_hosp[dataset$date>=day-7 & dataset$date <= day-1]))
  
  intensive_care_av7D = append(intensive_care_av7D,mean(dataset$intensive_care[dataset$date>=day-7 & dataset$date <= day-1]))
  
  hospitalized_with_symptoms_av7D = append(hospitalized_with_symptoms_av7D,mean(dataset$hospitalized_with_symptoms[dataset$date>=day-7 & dataset$date <= day-1]))
  
  day = day+1
}
```


#### Data by week
```{r}
start = as.Date("2020-12-06")
end = as.Date("2021-07-05")

intensive_care_week = c() 
intensive_care_weekH8 = c() 

newpos_week = c()
newpos_weekH8 = c()

hospitalized_with_symptoms_week = c()
hospitalized_with_symptoms_avg8H = c()

day = start
while(day <= end){
   
  newpos_week = append(newpos_week, mean(dataset$new_positives[dataset$date>=day-7 & dataset$date <= day-1]))
  
  intensive_care_weekH8 = append(intensive_care_weekH8, mean(dataset$intensive_care[dataset$date>=day & dataset$date <= day+6]))
  
  hospitalized_with_symptoms_week = append(hospitalized_with_symptoms_week,mean(dataset$hospitalized_with_symptoms[dataset$date>=day-7 & dataset$date <= day-1]))
  
day = day+7
  
}

data7 = data.frame(seq(1,31,1),intensive_care_weekH8,newpos_week,hospitalized_with_symptoms_week)
names(data7) = c("week","intensive_care_weekH8","newpos_week","hospitalized_with_symptoms_week")

```


#### total_hosp and intensive_care 7 days ahead

```{r}
dataPlus7 = subset(dataset, date>=start+7 & date<=end+7)
dataPlus7 = subset(dataPlus7,select=c(total_hosp, hospitalized_with_symptoms, intensive_care, deaths_per_day))
names(dataPlus7)=c("hospH8","symptomsH8","intcareH8", "deathsH8")
```

#### Turning colors into factors

```{r}
colors$colorLabel=factor(as.factor(colors$colorLabel),levels = c("Bianca","Gialla","Arancione","Rossa"))
names(colors) = c("date","color")
```

#### Select the time interval of interest

```{r}
dataset = subset(dataset, date>=start & date<=end)
colors = subset(colors, date>=start & date<=end)
```

#### Putting all together

```{r}
total = merge(dataset,colors,by="date")
total = cbind(total, dataPlus7, newpos_av7D, hosp_av7D, hospitalized_with_symptoms_av7D, intensive_care_av7D)
rownames(total) = NULL
```

#### Remove useless columns

```{r}
total=subset(total, select=c(date,newpos_av7D,hospitalized_with_symptoms_av7D,intensive_care_av7D,color,deathsH8))
```

###STANDARDIZE QUANTITATIVE COLUMNS+ KEEP NON STANDARDIZED NEW_POS_AV7D
```{r}
newpos_av7D_not_scaled=total$newpos_av7D
total = cbind(total,newpos_av7D_not_scaled)
total$newpos_av7DSCALED=scale(total$newpos_av7D)
total$hospitalized_with_symptoms_av7DSCALED=scale(total$hospitalized_with_symptoms_av7D)
total$intensive_care_av7DSCALED=scale(total$intensive_care_av7D)
total$deathsH8SCALED=scale(total$deathsH8)
```



#### Save 
```{r}
saveRDS(total,"dataset.rds")
saveRDS(data7,"weekly.rds")

```
