---
title: "ARl"
output: html_document
date: "2022-07-15"
---

# AR(1) of New positives

## Load Data

```{r}
rm(list=ls())
library(rjags)
dataset = readRDS("dataset.rds")
```

### Plot time series for new_positives covariate and newpos_av7D covariates

```{r}
new_positives.ts = ts(dataset$new_positives, frequency = 1)
plot(new_positives.ts )
newPos_av7D.ts = ts(dataset$newpos_av7D, frequency = 1)
plot(newPos_av7D.ts)
```

## Prior

$$ 
\begin{split}
& \alpha \sim {\cal Unif}(-1, 1) \\
& \mu \sim {\cal N}(0, 1000) \\
& \tau=\frac{1}{\sigma^2} \sim \mathcal{G}(0.001,0.001) 
\end{split}
$$


## Jags Model Formulation

For the prior, we assume that alpha and mu are following normal distributions and tau is following a gaussian distribution. 

```{r}
modelAR.string <-"model {
  ## parameters: alpha,tau,m0
  #likeliohood 
  mu[1] <- Y[1]
  for (i in 2:N) {
    Y[i] ~ dnorm(mu[i],tau)
    Yp[i] ~ dnorm(mu[i],tau) # prediction
    mu[i] <- m0 + alpha*Y[i-1]
  }
  sigma2<-1/tau
  
  #uninformative prior 
  alpha ~ dunif(-1,1)
  tau ~ dgamma(1.0E-3, 1.0E-3)
  m0 ~ dnorm(0.0, 1.0E-4)
}"
```


## Run the chain

We want to predict the number of new positives cases using the number of new positives at time t-1 (using the new_av7D variable).
```{r}
line_data =  list("Y" = dataset$newpos_av7D, "N"=length(dataset$newpos_av7D))
jagsAR = jags.model(textConnection(modelAR.string),
                    data=line_data,
                    n.chains = 2,
                    n.adapt = 300)
update(jagsAR,5000)
#We estimate the posterior of the two parameters alpha and sigma2 using MCMC technique.
outputmcmcAR=coda.samples(jagsAR,
                        c('alpha',"sigma2","mu","Yp"),
                        n.iter=10000, progress.bar="none")
```
### Summary
```{r}
#summary(outputmcmcAR)
```

## Trace Plots $\alpha$
```{r}
alpha.mcmc=outputmcmcAR[,c("alpha")]
plot(alpha.mcmc,main="posterior alpha")
```

## Trace Plots $\sigma^2$
```{r}
sigma.mcmc=outputmcmcAR[,c("sigma2")]
plot(sigma.mcmc,main="posterior sigma^2")
```

## True Data (red) and Sample Estimates (green) With Corresponding Confidence Intervals


```{r}
suAR=summary(outputmcmcAR)
n = length(dataset$newpos_av7D)
t=seq(1,n)
yp=suAR$statistics[1:n-1]
q1=suAR$quantiles[,1][1:n-1]
q2=suAR$quantiles[,5][1:n-1]
muest=suAR$statistics[1:n]
plot(dataset$date,
     dataset$newpos_av7D,
     col="red",
     ylab="newpos_av7D",
     main="newpos_av7D",
     xlab="time")
lines(dataset$date,q1[1:n])
lines(dataset$date,q2[1:n])
lines(dataset$date,muest,type="p",col="green")
```


```{r}
autocorr.plot(alpha.mcmc[2], main="alpha ACF")
autocorr.plot(sigma.mcmc[2], main="sigma2 ACF")
```
The first line shows the correlation of the parameter with itself (so it is always one). From the plots, we can observe that we don't have highly autocorrelated parameters because the vertical bars drop quickly to almost zero.
